---
title: "EPI590R Final Project"
author: "Alexis Barrett"
format:
  html: 
    theme: quartz
execute: 
  eval: true
  echo: false
  error: true
---

```{r}
#| output: false

#Setting up data and libraries
library(tidyverse)
library(gtsummary)
library(here)
library(broom)
library(gt)

working_dir <- here::here("data")
trashwheel <- read.csv(here("data", "trashwheel.csv"))

```
### Data Source:
The data is Trash Wheel Collection Data from the Mr. Trash Wheel Baltimore Healthy Harbor initiative (https://www.waterfrontpartnership.org/healthy-harbor-initiative).

Mr. Trash Wheel is a semi-autonomous trash interceptor that is placed at the end of a river, stream or other outfall. Far too lazy to chase trash around the ocean, Mr. Trash Wheel stays put and waits for the waste to flow to him. Sustainably powered and built to withstand the biggest storms, Mr. Trash Wheel uses a unique blend of solar and hydro power to pull hundreds of tons of trash out of the water each year.

The collection data for the Trash Wheels in Baltimore is obtained by sampling the material in each dumpster that is removed for disposal.

	
## Table

```{r}
#| label: tbl-one
#| tbl-cap: "Trash Wheel Collection Data"

table1 <- tbl_summary(
	trashwheel,
	by = Name,
	include = c(
		PlasticBottles, Polystyrene, CigaretteButts, GlassBottles,
		PlasticBags, Wrappers, SportsBalls
	),
	label = list(
		PlasticBottles ~ "Plastic Bottles",
		Polystyrene ~ "Polystyrene Items",
		CigaretteButts ~ "Cigarette Butts",
		GlassBottles ~ "Glass Bottles",
		PlasticBags ~ "Plastic Bags",
		Wrappers ~ "Wrappers",
		SportsBalls ~ "Sports Balls"
	),
	missing_text = "Missing"
) |>
	modify_header(label ~ "**Trash Collected**") |>
	modify_spanning_header(all_stat_cols() ~ "**Trash Wheel Name**") |>
	add_p(pvalue_fun = label_style_pvalue(digits = 2)) |>
	add_overall(col_label = "**Overall**") |>
	bold_labels()

table1
```

```{r}
# Save table as HTML using gt
gt_table <- as_gt(table1)

# Define the file path for the HTML file
html_file_path <- here("documents", "summary_table.html")

# Save the gt table as HTML
gtsave(gt_table, filename = html_file_path)
```




## Regression

```{r}
new_table_function <- function(model) {
	tbl_regression(
		model,
		label = list(
					HomesPowered ~ "Homes Powered",
					PlasticBottles ~ "Plastic Bottles Collected",
					Polystyrene ~ "Polystyrene Items Collected",
					CigaretteButts ~ "Cigarette Butts Collected",
					GlassBottles ~ "Glass Bottles Collected",
					PlasticBags ~ "Plastic Bags Collected",
					Wrappers ~ "Wrappers Collected",
					SportsBalls ~ "Sports Balls Collected"
		),
	)
}

linear_model <- lm(Weight ~ HomesPowered + PlasticBottles + Polystyrene + CigaretteButts + GlassBottles + PlasticBags + Wrappers + SportsBalls,
  data = trashwheel
)

```

```{r}
#| label: tbl-linear
#| tbl-cap: "Linear Regression Model"
new_table_function(linear_model)
```




## Figure

```{r}
#| label: fig-hist
#| fig-cap: "Homes Powered Over the Years"

figure1 <- ggplot(trashwheel, aes(x = as.factor(Year), y = HomesPowered)) +
	geom_bar(stat = "identity", fill = "skyblue") +
	labs(x = "Year", y = "Number of Homes Powered") +
	annotate("text", x = Inf, y = -Inf, label = "Note: Each ton of trash equals an average 500 kW of electricity. An average household will use 30 kW daily.", hjust = 1, vjust = -1, size = 3, color = "black") +
	theme_minimal() +  # Clean theme
	theme(axis.text.x = element_text(angle = 45, hjust = 1),
				panel.background = element_rect(fill = "white"),
    		plot.background = element_rect(fill = "white")
				)

figure1
```

```{r}
# Define the file path using here for saving the figure
file_path <- here("documents", "homes_powered_fig.png")

# Save the figure
ggsave(file_path, plot = figure1, width = 10, height = 6)
```




## InLine & Cross-Reference

```{r}
stats <- list( 
  n = n_distinct(trashwheel$Name), # Total trash wheels
  
  tot_trash = trashwheel %>%
    summarise(across(c(PlasticBottles, PlasticBags), \(x) sum(x, na.rm = TRUE))) %>%
    summarise(total_trash = sum(across(everything()), na.rm = TRUE)) %>%
    pull(total_trash), # Total trash collected
  
  tot_year = trashwheel %>%
    summarise(min_year = min(Year, na.rm = TRUE), 
              max_year = max(Year, na.rm = TRUE)) %>%
    summarise(total_years = max_year - min_year + 1) %>%
    pull(total_years) # Total number of years
)

professor_cig <- inline_text(table1, variable = "CigaretteButts", column = "stat_4")

gwynnda_cig <- inline_text(table1, variable = "CigaretteButts", column = "stat_2")

# Format number with commas
formatted_stats <- list(
  total_trash = format(stats$tot_trash, big.mark = ",")
)
```

-   @fig-hist illustrates the number of homes powered each year, with a note highlighting that each ton of trash generates an average of 500 kW of electricity, sufficient for an average household's daily consumption of 30 kW.
-   @tbl-one displays the types and quantities of trash collected, categorized by trash wheel name, including statistics and overall results, with missing values labeled as 'Missing' and p-values highlighted.
- @tbl-linear contains results from the linear regression.
-   There were `r stats$n` Trash Wheels in the Baltimore Healthy Harbor initiative, with a total of `r formatted_stats$total_trash` pieces of plastic bottles and bags collected over `r stats$tot_year` years.
-   Professor Trash Wheels collected a greater proportion of cigarette butts (median (Q1,Q3): `r professor_cig`) compared to the Gwynnda Trash Wheels (median (Q1,Q3): `r gwynnda_cig`).
